Reparametrization of COM-Poisson Regression Models with Applications in the Analysis of Experimental Data

### 2017-11-17 article.Rnw
Add an alternative graph to show bias results and fixed little things

library(knitr)
library(xtable)
options(digits = 3, OutDec = ".",
        xtable.caption.placement = "top",
        xtable.booktabs = TRUE,
        xtable.sanitize.text.function = identity,
        xtable.size = "small",
        xtable.math.style.negative = TRUE)
opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    echo = FALSE,
    results = "hide",
    fig.width = 7,
    fig.height = 5,
    out.width = "1\\textwidth",
    fig.align = "center",
    fig.pos = "!ht",
    # dev = "tikz"
    dev.args = list(family = "Palatino")
    )

@@

The results presented in \autoref{fig:approx-plot}(a) support the
proposed reparametrization. Replacing $\lambda$ and $\nu$ as function of
$\mu$ and $\phi$ in \autoref{eqn:pmf-cmp}, the reparametrized
distribution takes the form
\begin{equation}
  \label{eqn:pmf-cmpmu}
  \Pr(Y=y \mid \mu, \phi) =
  \left ( \mu +\frac{ e^\phi-1}{2e^\phi} \right )^{ye^\phi}
  \frac{(y!)^{-e^\phi}}{Z(\mu, \phi)},
  \qquad y = 0, 1, 2, \ldots\,,
\end{equation} where $\mu > 0$. We denotes this distribution as
COM-Poisson$_\mu$. In \autoref{fig:pmf-cmp} we show the shapes of
COM-Poisson$_\mu$ distribution.

@@

Parameter estimation requires numerical maximization of
\autoref{eqn:ll-rcmp}. Since derivatives of $\ell$ cannot be obtained in
closed forms, we adopted the BFGS algorithm~\cite{Nocedal1995} as
implemented in the function \texttt{optim()} for the statistical
software \texttt{R}. Standard errors for regression coefficients are
obtained based on the observed information matrix
$\mathcal{I}(\bm{\theta})$, where
$\mathcal{I}(\bm{\theta}) = -\mathcal{H}(\bm{\theta})$ (Hessian matrix)
is computed numerically. Confidence intervals for $\hat{\mu}_i$ are
obtained by using the delta method~\citep[p. 89]{Pawitan2001}.

@@

<<bias-plot, cache=TRUE, fig.height=5, fig.width=8, fig.cap="Distributions of standardized bias (gray points) and average with confidence intervals (black segments) by differents sample sizes and dispersion levels.">>=

@@

xyplot(param ~ bias | phi, groups = n, data = bias,
       panel = panel.superpose,
       layout = c(4, 1),
       key = key,
       strip = strip.custom(factor.levels = fl),
       scales = list(y = list(labels = yl)),
       ylab = "",
       xlab = "Standardized Bias",
       jitter.y = TRUE,
       factor = 0.1,
       gap = 0.3,
       alpha = 0.6,
       col = "gray80",
       cex = 0.7,
       panel.groups = function(x, y, group.number,
                               subscripts, gap, ...){
           noise <- centfac(factor(levels(bias$n)), space = gap)
           noise <- sort(noise)
           panel.xyplot(x, y + noise[group.number], ...)
       }) +
    as.layer(
        segplot(param ~ bias[, "lwr"] + bias[, "upr"] | phi,
                centers = bias[, "fit"],
                data = cidata,
                draw = FALSE,
                horizontal = TRUE,
                layout = c(4, 1),
                groups = n, gap = 0.3,
                key = key,
                pch = 21:24,
                cex = 0.7,
                lwd = 2,
                panel = function(...) {
                    panel.groups.segplot(...)
                    panel.abline(v = 0, lty = 2)
                    panel.abline(h = 1:5, col = "lightgray", lty = 2)
                })
    )
@

@@

<<coverage-plot, fig.height=4, fig.width=9, fig.cap="Coverage rate based on confidence intervals obtained by quadratic approximation for differents sample sizes and dispersion levels.">>=

@@

\begin{table}[ht]
\centering
\caption{Empirical correlations between $\hat{\phi}$ and $\hat{\bm{\beta}}$ for the two parametrizations of COM-Poisson model fit to equidispersed data.}
\label{tab:corr-cotton}
\begingroup\small
\begin{tabular}{lrrrrrr}
  \toprule
<<results="asis">>=

>>>>

library(knitr)
library(xtable)
options(digits = 3, OutDec = ".",
        xtable.caption.placement = "top",
        xtable.booktabs = TRUE,
        xtable.sanitize.text.function = identity,
        xtable.size = "small",
        xtable.math.style.negative = TRUE)
opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    echo = FALSE,
    results = "hide",
    fig.width = 7,
    fig.height = 5,
    out.width = "1\\textwidth",
    fig.align = "center",
    fig.pos = "!htp",
    # dev = "tikz"
    dev.args = list(family = "Palatino")
    )

@@

The results presented in \autoref{fig:approx-plot}(a) support the
proposed reparametrization. Replacing $\lambda$ and $\nu$ as function of
$\mu$ and $\phi$ in \autoref{eqn:pmf-cmp}, the reparametrized
distribution takes the form
\begin{equation}
  \label{eqn:pmf-cmpmu}
  \Pr(Y=y \mid \mu, \phi) =
  \left ( \mu +\frac{ e^\phi-1}{2e^\phi} \right )^{ye^\phi}
  \frac{(y!)^{-e^\phi}}{Z(\mu, \phi)},
  \qquad y = 0, 1, 2, \ldots\,,
\end{equation} where $\mu > 0$. We denote this distribution as
COM-Poisson$_\mu$. In \autoref{fig:pmf-cmp} we show the shapes of
COM-Poisson$_\mu$ distribution.

@@

Parameter estimation requires the numerical maximization of
\autoref{eqn:ll-rcmp}. Since the derivatives of $\ell$ cannot be
obtained in closed forms, we adopted the \texttt{BFGS}
algorithm~\cite{Nocedal1995} as implemented in the function
\texttt{optim()} for the statistical software \texttt{R}. Standard
errors for the regression coefficients are obtained based on the
observed information matrix $\mathcal{I}(\bm{\theta})$, where
$\mathcal{I}(\bm{\theta}) = -\mathcal{H}(\bm{\theta})$ (Hessian matrix)
is computed numerically. Confidence intervals for $\hat{\mu}_i$ are
obtained by using the delta method~\citep[p. 89]{Pawitan2001}.

@@

<<bias-plot, cache=TRUE, fig.height=6.5, fig.width=8.5, fig.cap="Distributions of standardized bias (gray points) and average with confidence intervals (black segments) by different sample sizes and dispersion levels.">>=

@@

# Alternative graph (figure with all point is very large in size memory)
bwplot(param ~ bias | phi, groups = n, data = bias,
       pch = "|",
       grid = TRUE,
       as.table = TRUE,
       layout = c(4, 1),
       xlab = "Standardized Bias",
       fill = "gray80",
       col = "red",
       strip = strip.custom(factor.levels = fl),
       panel = panel.superpose,
       scales = list(y = list(labels = yl)),
       box.width = 0.05,
       whisker.width = 0.4,
       key = key,
       par.settings = list(
           plot.symbol = list(pch = 20, cex = 0.5, col = "gray70"),
           box.rectangle = list(col = "gray70"),
           box.umbrella=list(col = "gray70", lwd = 2)
       ),
       panel.groups = function(x, y, group.number, ...) {
           my.panel.bwplot(x, (y - 0.05) + (group.number - 2.5) / 5, ...)
           panel.abline(v = 0, lty = 2)
       }) +
    segplot(param ~ bias[, "lwr"] + bias[, "upr"] | phi,
            centers = bias[, "fit"],
            data = cidata,
            draw = FALSE,
            horizontal = TRUE,
            layout = c(4, 1),
            groups = n, gap = 0.3,
            key = key,
            pch = 21:24,
            cex = 0.7,
            lwd = 2,
            panel = function(...) {
                panel.groups.segplot(...)
                panel.abline(v = 0, lty = 2)
                panel.abline(h = 1:5, col = "lightgray", lty = 2)
            })

# xyplot(param ~ bias | phi, groups = n, data = bias,
#        panel = panel.superpose,
#        layout = c(4, 1),
#        key = key,
#        strip = strip.custom(factor.levels = fl),
#        scales = list(y = list(labels = yl)),
#        ylab = "",
#        xlab = "Standardized Bias",
#        jitter.y = TRUE,
#        factor = 0.1,
#        gap = 0.3,
#        alpha = 0.6,
#        col = "gray80",
#        cex = 0.7,
#        panel.groups = function(x, y, group.number,
#                                subscripts, gap, ...){
#            noise <- centfac(factor(levels(bias$n)), space = gap)
#            noise <- sort(noise)
#            panel.xyplot(x, y + noise[group.number], ...)
#        }) +
#     as.layer(
#         segplot(param ~ bias[, "lwr"] + bias[, "upr"] | phi,
#                 centers = bias[, "fit"],
#                 data = cidata,
#                 draw = FALSE,
#                 horizontal = TRUE,
#                 layout = c(4, 1),
#                 groups = n, gap = 0.3,
#                 key = key,
#                 pch = 21:24,
#                 cex = 0.7,
#                 lwd = 2,
#                 panel = function(...) {
#                     panel.groups.segplot(...)
#                     panel.abline(v = 0, lty = 2)
#                     panel.abline(h = 1:5, col = "lightgray", lty = 2)
#                 })
#     )

@@

<<coverage-plot, fig.height=4, fig.width=9, fig.cap="Coverage rate based on confidence intervals obtained by quadratic approximation for different sample sizes and dispersion levels.">>=

@@

\begin{table}[ht]
\centering
\caption{Empirical correlations between $\hat{\phi}$ and $\hat{\bm{\beta}}$ for the two parametrizations of COM-Poisson model fit to underdispersed data.}
\label{tab:corr-cotton}
\begingroup\small
\begin{tabular}{lrrrrrr}
  \toprule
<<results="asis">>=
